Self-assembly of colloid-cholesteric composites: A route to switchable optical materials

### 2013-10-02 papers/2012-sandwich/CholColloids.tex
Minor improvements to text
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2165 40876441-3496-0410-a613-a2fca35b2ae2

{\bf Colloidal particles dispersed in liquid crystals can form new materials with tunable elastic and electro-optic properties~\cite{stark}. In a host comprising  a periodic `blue phase' chiral liquid crystal~\cite{mermin}, the particles should template into colloidal crystals \cite{miha} with potential uses in photonics, metamaterials, and transformational optics~\cite{lavrentovich}. 
Here we show by computer simulation that colloid/cholesteric mixtures can give rise to a much wider range of soft-solid composites than previously anticipated. These include regular crystals, glasses, percolating gels, isolated clusters, twisted rings and undulating colloidal ropes. The final structure can be tuned through the particle concentration, and by varying the surface interactions of the cholesteric host with both the colloids and any confining walls. 
Remarkably, we find that many of these new materials are metastable: two or more structures can arise under identical thermodynamic conditions. The observed structure depends not only on the formulation protocol, but also on the field history.  
This new class of soft materials should thus be relevant to the design of switchable, multistable devices for optical technologies including, for example, e-paper~\cite{epaper}.}

When spherical colloidal particles are mixed into a nematic liquid crystal, 
they disrupt the orientational order in the fluid and create defects (disclination lines) in the nematic close to their surfaces. To minimize their free energy cost, it is generally advantageous for defects to be shared:
therefore the colloidal particles have a generic tendency to aggregate. Such aggregates may further 
self-assemble into lines~\cite{wiresmiha}, 2D crystals~\cite{zumer}, 
planar structures~\cite{tanaka}, or 3D amorphous glasses~\cite{tiffany}.
Besides being of fundamental interest to materials science, these
structures have tunable elastic and optical properties~\cite{stark}. Hence 
they offer exciting prospects for applications as biosensors~\cite{abbott}, or
as new-generation display devices as they can exploit the bistability
inherent in cholesteric devices~\cite{colloiddevice,choldevice}.

@@

For a single colloidal particle in a nematic fluid, $w$ controls the formation of a local topological defect at the particle surface~\cite{stark}, with weak distortion of the nearby fluid at small $w$. In the colloid-BP system, $w$ also plays a determining role in the final composite structure. When $w$ is small (Fig.~ 1A,B), the attraction of particles to the disclination lattice causes this to become covered by particles, with little disruption of its long-range order. Thus a templated colloidal crystal is created, as first reported in~\cite{miha}. Interestingly, the weak but finite anchoring apparently creates an elastic force between the colloids which leads to the formation of small colloidal lines within disclinations, especially for dilute samples (Fig. 1A). It seems likely that polymer-stabilised  blue phases~\cite{kikuchi} work by a similar principle to this small $w$ case, with a weak segregation of monomers towards the defect regions.

@@

Fig.~2A-D shows the case of normal wall anchoring on varying $\phi$ and $w$. We quench from the isotropic to the ordered phase as was done in Fig.~1E-H, and again find a metastable, amorphous disclination network (see Supporting Information). The resulting colloid-cholesteric composites are somewhat similar to those seen in the bulk: for small $w$ the particles are templated by the network (Fig.~2A,B), whereas for large $w$ we observe isolated clusters (Fig.~2C) and percolated colloidal gels (Fig.~2D), for $\phi = 1\%$ and $2\%$ respectively. There are, however, significant effects of the confining walls.
First, the normal anchoring at the wall recruits particles there,  giving (for the chosen $h$) a density enhancement of the colloids at both walls and also in the midplane (see side-views in Fig.~2A-D). Second, the clusters formed at large $w$ are themselves anisotropic with an oblate habit. These are similarly to those reported experimentally for large colloids in a cholesteric (but non-BP) host~\cite{niek}.

@@

An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting again from an amorphous blue phase,
this time at a higher chirality. Dispersing colloids in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A). Applying a high enough field along the $z$ 
direction leads to a switching to a new phase, where colloidal arrays fit 
within a network with distorted hexagonal cells (Fig.~3B). Remarkably, upon 
field removal this honeycomb structure does not find its way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Fig. S7), with residual anisotropy along the field switching 
direction.
The resulting multistability is indeed a must-have feature for new generation 
switchable electrooptic devices for use in energy-saving electronics-free
appliances, smart glass, and e-paper to cite just a few examples.

>>>>

{\bf Colloidal particles dispersed in liquid crystals can form new materials with tunable elastic and electro-optic properties~\cite{stark}. In a host comprising  a periodic `blue phase' chiral liquid crystal~\cite{mermin}, the particles should template into colloidal crystals \cite{miha} with potential uses in photonics, metamaterials, and transformational optics~\cite{lavrentovich}. 
Here we show by computer simulation that colloid/cholesteric mixtures can give rise to a much wider range of soft-solid composites than previously anticipated. These include regular crystals, glasses, percolating gels, isolated clusters, twisted rings and undulating colloidal ropes. The final structure can be tuned through the particle concentration, and by varying the surface interactions of the cholesteric host with both the colloids and any confining walls. 
Remarkably, we find that many of these new materials are metastable: two or more structures can arise under identical thermodynamic conditions. The observed structure depends not only on the formulation protocol,
but also on the history of an applied electric field.  
This new class of soft materials should thus be relevant to the design of switchable, multistable devices for optical technologies including, for example, e-paper~\cite{epaper}.}

When spherical colloidal particles are mixed into a nematic liquid crystal, 
they disrupt the orientational order in the fluid and create defects (disclination lines) in the nematic close to their surfaces. To minimize the free energy cost, it is generally advantageous for defects to be shared:
therefore the colloidal particles have a generic tendency to aggregate. Such aggregates may further 
self-assemble into lines~\cite{wiresmiha}, 2D crystals~\cite{zumer}, 
planar structures~\cite{tanaka}, or 3D amorphous glasses~\cite{tiffany}.
Besides being of fundamental interest to materials science, these
structures have tunable elastic and optical properties~\cite{stark}. Hence 
they offer exciting prospects for applications as biosensors~\cite{abbott}, or
as new-generation display devices as they can exploit the bistability
inherent in liquid crystal devices~\cite{colloiddevice,choldevice}.

@@

For a single colloidal particle in a nematic fluid, $w$ controls the formation of a local topological defect at the particle surface~\cite{stark}, with weak distortion of the nearby fluid at small $w$. In the colloid-BP system, $w$ also plays a determining role in the final composite structure. When $w$ is small (Fig.~ 1A,B), the attraction of particles to the disclination lattice causes
 the lattice to become covered by particles, with little disruption of its
 long-range order. Thus a templated colloidal crystal is created, as first reported in~\cite{miha}. Interestingly, the weak but finite anchoring apparently creates an elastic force between the colloids which leads to the formation of small colloidal lines within disclinations, especially for dilute samples (Fig. 1A). It seems likely that polymer-stabilised  blue phases~\cite{kikuchi} work by a similar principle to this small $w$ case, with a weak segregation of monomers towards the defect regions.

@@

Fig.~2A-D shows the case of normal wall anchoring on varying $\phi$ and $w$. We quench from the isotropic to the ordered phase as was done in Fig.~1E-H, and again find a metastable, amorphous disclination network (see Supporting Information). The resulting colloid-cholesteric composites are somewhat similar to those seen in the bulk: for small $w$ the particles are templated by the network (Fig.~2A,B), whereas for large $w$ we observe isolated clusters (Fig.~2C) and percolated colloidal gels (Fig.~2D), for $\phi = 1\%$ and $4\%$ respectively. There are, however, significant effects of the confining walls.
First, the normal anchoring at the wall recruits particles there,  giving (for the chosen $h$) a density enhancement of the colloids at both walls and also in the midplane (see side-views in Fig.~2A-D). Second, the clusters formed at large $w$ are themselves anisotropic with an oblate habit. These clusters
are similar to those reported experimentally for large colloids in a
cholesteric (but non-BP) host~\cite{niek}.

@@

An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting again from an amorphous blue phase,
this time at a higher chirality. Dispersing colloids in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A). Applying a high enough field along the $z$ 
direction leads to a switching to a new phase, where colloidal arrays fit 
within a network with distorted hexagonal cells (Fig.~3B). Remarkably, upon 
field removal this honeycomb structure does not find its way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Fig. S7), with residual anisotropy along the field switching 
direction.
The resulting multistability is indeed a prerequisite for new generation 
switchable electrooptic devices for use in energy-saving electronics-free
appliances, smart glass, and e-paper to cite just a few examples.
